SU($n$) spin models near the Heisenberg critical point

### 2020-03-19 collective_sun/sun_perturbation.tex
emphasized that *spatial* symmetries simplify the multi-body eigenvalue problem broke off the subsubsection about evaluating multi-spin operators into its own subsection

\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  This projector essentially
picks off the part of $X_{\m w}$ that is strictly off-diagonal with
respect to the ground- and excited-state manifolds $\E_0$ and
$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
states of definite excitation energy when acting on PS states $\M_0$,
we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\check{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but symmetries of
$\hat{\m h}$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

@@

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul{\M}_M\setminus\M_0$.  If
the coupling between $\M_0$ and its complement is weak compared to the
spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul{\M}_M$, then $\ul{\M}_M$ to
$\ul{\M}_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory analogous to first-order perturbation theory on
$\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul{\M}_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
simulations restricted to $\ul{\M}_{r_{\t{max}}}$ can be benchmarked
by examining the maximum populations within all $\M_r$ for
$r\le r_{\t{max}}$ throughout simulation.  As long as these
populations fall off with increasing $r$ and the maximal population
within large-$r$ manifolds $\M_r$ is small, one should expect such
simulations to faithfully capture (at least qualitatively) the
dynamical behavior of the spin system under consideration.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Translational invariance, isotropy, and other symmetries}
\label{sec:symmetries}

Solving the eigenvalue problem in \eqref{eq:multi_body_eig} to find
$M$-body perturbations $X_{\m w}$ that generate states of definite
interaction energy nominally requires diagonalizing the matrix
$\hat{\m h}$ with dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
interaction Hamiltonian $H_0$ and multi-body perturbation $X_{\m w}$
have any symmetries, such as translational invariance isotropy, then
we can use these symmetries to reduce the complexity of this
eigenvalue problem.  Translational invariance, for example, can reduce
this problem to that of diagonalizing a matrix with dimensions
$\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times
N^{M-1}$, and isotropy can further reduce the problem size by factors
that depend on the lattice geometry.  The complexity of diagonalizing
a matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a
reduction of $d$ by a factor of 2 yields practically significant gains
(a factor of $2^3=8$) in the runtime of diagonalization.  Here, we
provide a general prescription for exploiting the symmetries of $H_0$
and $X_{\m w}$ to simplify the eigenvalue problem in
\eqref{eq:multi_body_eig}.

Symmetries of $H_0$ and $X_{\m w}$ essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig} in a basis that obeys the symmetries
under consideration.  To this end, for any equivalence class
$s\in\S_M$ we define a uniform superposition $\ket{s}$ over all
choices $k\in s$, as well as the projector $\P_\S$ onto the space of
uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  We then expand the
coefficient vector $\v{\m w}$ in this symmetrized basis:
\begin{align}
  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \hat{\m h}_{\S_M}
  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \ket{s} \sp{
    \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
    - \tilde h_s \bra{s}},
\end{align}
where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
$a$-th element of $k$;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
is the equivalence class of $k_{a:p}$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Matrix elements of individual multi-spin operators}

The process of constructing and simplifying diagrams to compute the
coefficients $w\p{g}$ of the diagrammatic operator product expansion
in \eqref{eq:sym_prod_group} is systematic enough to be converted into
an algorithm for execution on a computer.  The last step necessary to
automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
the multi-spin operators $X\p{g}$ in a basis for the permutationally
symmetric manifold $\M_0$.  Without loss of generality, we consider a
single permutationally symmetric $M$-spin operator $X$; our task is
essentially to find coefficients of the expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b} \bk{a|X|b} \op{a}{b},
\end{align}
where $\set{\ket{a}},\set{\ket{b}}$ are complete bases for $\M_0$, and
the choice of $M$ spins on which $X$ acts is arbitrary.  In a system
of $N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
labeled by the occupation number of each single-spin state, namely
$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_j\ge 0$ and $\sum_ja_j=N$.
Written out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_{j=1}^n \ket{j}^{\otimes a_j},
  &&
  \C\p{a} \equiv \f{\p{\sum_j a_j}!}{\prod_ja_j!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
$S$ spins to $n$ states by $\A_n\p{S}$, such that the standard basis
for the permutationally symmetric manifold of $S$ spins is
$\set{\ket{a}:a\in\A_n\p{S}}$.  Identifying this standard basis allows
us to expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_j\le a_j$ and
$\p{a-\alpha}_j=a_j-\alpha_j$ for all $j$; and $\delta_{cd}=1$ if
$c=d$ and zero otherwise.  We sum over both $\alpha$ and $\beta$ above
merely to keep the expression symmetric with respect to transposition;
in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set
$\beta=b-a+\alpha$.  Note that, by slight abuse of notation, the
operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an
arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$
on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with
respect to the $M$-spin states $\ket\alpha,\ket\beta$.

>>>>

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$ (with respect to $H_0$).  This projector essentially
picks off the part of $X_{\m w}$ that is strictly off-diagonal with
respect to the ground- and excited-state manifolds $\E_0$ and
$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
states of definite excitation energy when acting on PS states $\M_0$,
we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ of the
interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
is
\begin{align}
  \p{\m h \circ_a \m w}_k
  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
\end{align}
where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
explicitly exclude terms with $p\in k$ that vanish due to the fact
that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
dimension-$M$ tensor $\tilde{\m h}$ with entries
\begin{align}
  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
\end{align}
for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
entries $\tilde h_k w_k$.  These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\check{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
symmetries of $H_0$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem.  We
discuss this reduction in Appendix \ref{sec:symmetries}.

@@

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul{\M}_M\setminus\M_0$.  If
the coupling between $\M_0$ and its complement is weak compared to the
spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul{\M}_M$, then $\ul{\M}_M$ to
$\ul{\M}_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory analogous to first-order perturbation theory on
$\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul{\M}_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
simulations restricted to $\ul{\M}_{r_{\t{max}}}$ can be benchmarked
by examining the maximum populations within all $\M_r$ for
$r\le r_{\t{max}}$ throughout simulation.  As long as
\begin{enumerate*}
\item these
populations fall off with increasing $r$,
\item and the maximal population within large-$r$ manifolds $\M_r$ is
  small,
\end{enumerate*}
one should expect such simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.

In order to project operators onto $\ul{\M}_{r_{\t{max}}}$, we need to
construct a basis for each $\M_r$ with $r\le r_{\t{max}}$.
\red{[TODO: finish section]}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Exploiting spatial symmetries}
\label{sec:symmetries}

Solving the eigenvalue problem in \eqref{eq:multi_body_eig} to find
$M$-body perturbations $X_{\m w}$ that generate states of definite
interaction energy nominally requires diagonalizing the matrix
$\hat{\m h}$ with dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
interaction Hamiltonian $H_0$ has any spatial symmetries, however,
then we can use these symmetries to reduce the complexity of this
eigenvalue problem.  Translational invariance, for example, can reduce
this problem to that of diagonalizing a matrix with dimensions
$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
and isotropy can further reduce the problem size by factors that
depend on the lattice geometry.  The complexity of diagonalizing a
matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
of $d$ by a factor of 2 yields practically significant gains (a factor
of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
general prescription for exploiting the spatial symmetries of $H_0$ to
simplify the eigenvalue problem in \eqref{eq:multi_body_eig}.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.

For simplicity, we first assume that the $M$-body operator $X_{\m w}$
obeys the same spatial symmetries as $H_0$.  For any equivalence class
$s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
choices $k\in s$, as well as the projector $\P_\S$ onto the space of
uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.  We then expand the
coefficient vector $\v{\m w}$ in this symmetrized basis:
\begin{align}
  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \hat{\m h}_{\S_M}
  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \ket{s} \sp{
    \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
    - \tilde h_s \bra{s}},
\end{align}
where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
$a$-th element of $k$;
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
is the equivalence class of $k_{a:p}$.  The reduced multi-body
eigenvalue problem is then
\begin{align}
  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
\end{align}
More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
obey the same spatial symmetries as $H_0$, these symmetries generally
endow $\check{\m h}$ with a block-diagonal structure. \red{[TODO:
  finish discussion/derivation]}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Matrix elements of individual multi-spin operators}

The process of constructing and simplifying diagrams to compute the
coefficients $w\p{g}$ of the diagrammatic operator product expansion
in \eqref{eq:sym_prod_group} is systematic enough to be converted into
an algorithm for execution on a computer.  The last step necessary to
automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
the multi-spin operators $X\p{g}$ in a basis for the permutationally
symmetric manifold $\M_0$.  Without loss of generality, we consider a
single permutationally symmetric $M$-spin operator $X$; our task is
essentially to find coefficients of the expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
\end{align}
where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
labeled by the occupation number of each single-spin state, namely
$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_\mu\ge 0$ and
$\sum_\mu a_\mu=N$.  Written out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
$N$ spins to $n$ states by $\A_n\p{N}$, such that the basis
$\set{\ket{a}:a\in\A_n\p{N}}$ spans permutationally symmetric manifold
of $N$ spins.  Identifying this basis allows us to expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
$\alpha$ and $\beta$ above merely to keep the expression symmetric
with respect to transposition; in practice, one can simply sum over
$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
abuse of notation, the operator $X$ on the left of
\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
(out of $N$), whereas the operator $X$ on the right of
\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
states $\ket\alpha,\ket\beta$.
